Response Letter for manuscript TSUSC-2018-12-0140 \\ ``Online EV Scheduling Algorithms for Adaptive Charging Networks with Global Peak Constraints'' \\ \vspace{4mm} \large by Bahram~Alinia, Mohammad~H.~Hajiesmaili, Zachary J. Lee, Noel Crespi, and Enrique Mallada

### 2019-08-27 TSC/R1/TSC.tex
BA - Update 2

\begin{abstract}
%			With the rapid growth of electric vehicles, their high demand charging jobs pose an immense challenge to the grid infrastructure. 
			This paper tackles online scheduling of electric vehicles (EVs) in an adaptive charging network (ACN) with local and global peak constraints. Given the aggregate charging demand of the EVs and the peak constraints of the ACN, it might be infeasible to fully charge all the EVs according to their charging demand. Two alternatives in such resource-limited scenarios are to maximize the social welfare by partially charging the EVs (fractional model) or selecting a subset of  EVs and fully charge them (integral model). 
%			This paper tackles both fractional and integral models. 
			The critical challenge is the need for online solution design since in practical scenario\revv{s} the scheduler has no information of future arrivals of EVs in a time-coupled underlying problem. 
			For the fractional model, we devise both offline and online algorithms. We prove that the offline algorithm is optimal. Using competitive ratio as the performance measure, we prove the online algorithm achieves a competitive ratio of 2. The integral model, however, is more challenging since the underlying problem is NP-hard due to 0/1 selection criteria of EVs. Hence, efficient solution design is challenging even in offline setting. We devise a low-complexity primal-dual scheduling algorithm that achieves a bounded approximation ratio. Built upon the offline approximate algorithm, we propose an online algorithm and analyze its competitive ratio in special cases. Extensive trace-driven experimental results show that the performance of the proposed online algorithms is close to the offline optimum, and outperform the existing solutions. 
%			\rev{In particular, the grid and electricity providers should be aware of the electricity peak.	It is known in electricity market that the spot price is a dominating factor in billing cost. Hence, the utility companies need to identify a global peak constraint for the cost management purpose. This paper studies the online EV charging scheduling problem in a utility that offers EV charging service in a set of CSs, each of which having limited output power (referred to as local peak constraint) while taking into account the global peak constraint. The scheduling problem in this scenario comes with unique challenges that is not tackled previously.}

@@

There are two main challenges in the design and implementation of scheduling algorithms for EVs satisfying the goals mentioned above. \rev{Firstly, the problem calls for online scheduling design.} In practice, EVs arrive to the CS in online fashion and the scheduler has no information about the arrival and demand of the future EVs. Secondly, the underlying optimization problem in integral model is NP-hard even in offline case \rev{(see Section~\ref{sec:integral})}. This is because the problem is a mixed integer linear problem and a ``time-expanded'' extension of knapsack problem which is known as a classic NP-hard problem. In this paper, we tackle the challenge of online design by following competitive algorithm design~\cite{borodin2005online} and the challenge of NP-hardness by pursuing approximation algorithm design~\cite{approx} and make the following contributions:

@@

\bt{This paper represents follow-on work to our previous work~\cite{alinia2018competitive}, where we study a simplified version of the problems studied in this paper in fractional revenue model. The competitive analysis in~\cite{alinia2018competitive} provide a competitive ratio of different algorithms when all EVs have the same maximum charging rate. Also, \cite{alinia2018competitive} does not study the integral revenue model. In this paper, extend the results and propose a 2-competitive online algorithm with no assumption on the input parameters. Also, the current study investigates both fractional and integral revenue models while considering global peak constraint in the formulation and addressing multiple charging station scenario. Last, we consider more realistic trace-driven experiments in this paper and compare the results to an existing real-world scheduling algorithm.}
%The simulation section also discusses on the positive and negative effects of slackness parameter in different scenarios.

@@

\subsubsection{Worst-case Analysis in Similar Scheduling Problems}
\label{sec:rel:comp}
\bt{
	Similar underlying scheduling problems with slightly different settings has been studied in the literature in both offline and online settings. In offline setting, the problem is more interested in integral revenue model where the problem becomes a combinatorial optimization problem and approximation algorithms have been used to find effective solutions. 
The performance of an approximation algorithm is determined by its \emph{approximation ratio} for offline 
algorithms. On the other hand, competitive online algorithms are used in the online setting and \emph{competitive ratio} is the performance metric which compare the algorithm's result to the offline optimal solution. 
Under integral revenue model, \cite{jain2015near} proposes an offline algorithm for scheduling of batch jobs in cloud computing which is similar to EV scheduling problem. It is assumed that all jobs are available to be processed at time 0. The authors propose a $\frac{C}{C-k}.\frac{s}{s-1}$-approximation algorithm where $C$ is the cloud capacity and $s$ is a ``slackness'' parameter. 
Similarly, \cite{yao2016real} tackles the offline integral problem and proposes a convex relaxation method to find a near-optimal solution. No theoretical bound is provided for the algorithm and the performance is examined by simulation results. In this paper, we propose \ics for the EV scheduling problem with slightly different settings in the constraint sets, and tackle the problem in online setting in which the full knowledge of the online parameters is not available a priori, and provide competitive ratio under specific scenarios. 
In case of fractional revenue model (for offline case), the scheduling problem is more straightforward to tackle since the underlying problem is linear. In this paper, we propose \fcs as an optimal solution for offline fractional case.

In the online scenario, the scheduling problem is fundamentally more challenging. In our recent study \cite{alinia2018competitive}, 
we proposed two online algorithms referred to as WFAIR and WRAND for EV scheduling problem. The proposed algorithms provide a competitive ratio of $2-1/U$ where $U$ is "scarcity level". However, \cite{alinia2018competitive} the result only holds when all EVs have the same maximum charging rate. Also, \cite{alinia2018competitive} does not study the integral revenue model. In this paper, we propose \focs as a 2-competitive online algorithm with no assumption on the input parameters. Also, the current study investigates both fractional and integral revenue model while considering total peak constraint in the formulation and addressing multiple charging station scenario. 
In another work~\cite{firstfit}, two simple and natural online algorithms called \textsc{FirstFit} and \textsc{EndFit} were developed and are proved to be 2-competitive. \focs is extension of these algorithms by taking into account the maximum charging rate of the EVs and multiple charging station scenario (see ``Remarks'' at Section IV). As recent studies, \cite{zheng2016online,deng2016whether} provide online algorithms for EV scheduling problem, however, they do not take into account global peak constraint in the underlying problem.

Another direction is to tackle online scenario under integral revenue model. In this setting, the scheduling problem becomes fundamentally more challenging. We note that the integral scheduling problem is strongly NP-Hard even in the offline case \cite{de2018complexity}. Due to combined online and combinatorial challenges, there are very limited studies in this category. \cite{lucier2013efficient} provides an online algorithm without considering the maximum processing rate of the jobs where its competitive ratio can be arbitrary bad depending on ``slackness'' parameter'. Our work in Section~\ref{sec:integral} also takes into account a notion of slackness parameter in our proposed online solution, and in addition, considers \iocs for multiple station scenario and maximum charging rate.
}

@@

\revv{
\subsection{Optimal Offline Design}
\label{sec:fopt}				
We refer the proposed algorithm as the \fcs and summarize it as Algorithm~\ref{alg:optimal-off}.
The \fcs works in two phases. In the first phase (Section~\ref{sec:ph1}), the algorithm decides on the amount of resource to be allocated to each EV within its availability window and reserves resources accordingly. In this phase, the details of allocation is not known. The actual resource allocation is done in the second phase (Section~\ref{sec:ph2}) by setting variables $y_i^t$.

@@

		\vspace{-3mm}
		\begin{algorithm}%\small%1
			\footnotesize
			\caption{\ics}
			\label{alg:ics}
			\DontPrintSemicolon 
%			\KwIn{$n$ EVs with $a_i, d_i, v_i, D_i,$ and $k_i$ associated with each EV $i$, $m$ CSs, local and global peak constraints $p_j, j=1,\dots ,m$, and $p^\mathsf{total}$}

@@

\vspace{-4mm}
\begin{algorithm}%\small%1
\footnotesize
\caption{\iocs: $\forall t\in \{1,2,\dots ,T\}$}
\label{alg:iocs}
\DontPrintSemicolon 
%\KwIn{$n$ EVs to arrive on the fly, number of CSs $m$, global peak constraint $p^{\textrm{total}}$, local peak constraints $p_j, j=1,\dots ,m$}

>>>>

\begin{abstract}
%			With the rapid growth of electric vehicles, their high demand charging jobs pose an immense challenge to the grid infrastructure. 
			This paper tackles online scheduling of electric vehicles (EVs) in an adaptive charging network (ACN) with local and global peak constraints. Given the aggregate charging demand of the EVs and the peak constraints of the ACN, it might be infeasible to fully charge all the EVs according to their charging demand. Two alternatives in such resource-limited scenarios are to maximize the social welfare by partially charging the EVs (fractional model) or selecting a subset of  EVs and fully charge them (integral model). 
%			This paper tackles both fractional and integral models. 
			The critical challenge is the need for online solution design since in practical scenario\revv{s} the scheduler has no information of future arrivals of EVs in a time-coupled underlying problem. 
			For the fractional model, we devise both offline and online algorithms. We prove that the offline algorithm is optimal. Using competitive ratio as the performance measure, we prove the online algorithm achieves a competitive ratio of 2. The integral model, however, is more challenging since the underlying problem is \bt{strongly} NP-hard due to 0/1 selection criteria of EVs. Hence, efficient solution design is challenging even in offline setting. We devise a low-complexity primal-dual scheduling algorithm that achieves a bounded approximation ratio. Built upon the offline approximate algorithm, we propose an online algorithm and analyze its competitive ratio in special cases. Extensive trace-driven experimental results show that the performance of the proposed online algorithms is close to the offline optimum, and outperform the existing solutions. 
%			\rev{In particular, the grid and electricity providers should be aware of the electricity peak.	It is known in electricity market that the spot price is a dominating factor in billing cost. Hence, the utility companies need to identify a global peak constraint for the cost management purpose. This paper studies the online EV charging scheduling problem in a utility that offers EV charging service in a set of CSs, each of which having limited output power (referred to as local peak constraint) while taking into account the global peak constraint. The scheduling problem in this scenario comes with unique challenges that is not tackled previously.}

@@

There are two main challenges in the design and implementation of scheduling algorithms for EVs satisfying the goals mentioned above. \rev{Firstly, the problem calls for online scheduling design.} In practice, EVs arrive to the CS in online fashion and the scheduler has no information about the arrival and demand of the future EVs. Secondly, the underlying optimization problem in integral model is \bt{strongly} NP-hard even in offline case \rev{(see Section~\ref{sec:integral})}. This is because the problem is a mixed integer linear problem and a ``time-expanded'' extension of knapsack problem which is known as a classic NP-hard problem. In this paper, we tackle the challenge of online design by following competitive algorithm design~\cite{borodin2005online} and the challenge of NP-hardness by pursuing approximation algorithm design~\cite{approx} and make the following contributions:

@@

\bt{This paper represents follow-up work to our previous study~\cite{alinia2018competitive}, where we address a simplified version of the problems studied in this paper in fractional revenue model. The competitive analysis in~\cite{alinia2018competitive} is done under the assumption that all EVs have the same maximum charging rate. Also, \cite{alinia2018competitive} does not study the integral revenue model. In this paper, we extend the results and propose a 2-competitive online algorithm with no assumption on the input parameters. Also, the current study investigates both fractional and integral revenue models while considering global peak constraint in the formulation and addressing multiple charging station scenario. Last, we consider more realistic trace-driven experiments in this paper and compare the results to an existing real-world scheduling algorithm.}
%The simulation section also discusses on the positive and negative effects of slackness parameter in different scenarios.

@@

\subsubsection{Worst-case Analysis in Similar Scheduling Problems}
\label{sec:rel:comp}
\bt{
	Similar underlying scheduling problems with slightly different settings have been studied in the literature in both offline and online settings. In offline setting, the problem is more interested in integral revenue model where the problem becomes a combinatorial optimization problem and approximation algorithms have been used to find effective solutions. 
The performance of an approximation algorithm is determined by its \emph{approximation ratio} for offline 
algorithms. On the other hand, competitive online algorithms are used in the online setting and \emph{competitive ratio} is the performance metric which compares the algorithm's result to the offline optimal solution.

\textbf{Offline integral model}: Under integral revenue model, \cite{jain2015near} proposes an offline algorithm for scheduling of batch jobs in cloud computing which is similar to EV scheduling problem. It is assumed that all jobs are available to be processed at time 0. The authors propose a $\frac{C}{C-k}.\frac{s}{s-1}$-approximation algorithm where $C$ is the cloud capacity and $s$ is the ``slackness parameter'' (See section \ref{sec:acn} for the definition). 
Similarly, \cite{yao2016real} tackles the offline integral problem and proposes a convex relaxation method to find a near-optimal solution. No theoretical bound is provided for the algorithm and the performance is examined by simulation results. In this paper, we propose \ics for the EV scheduling problem with slightly different settings in the constraint sets, and tackle the problem in online setting in which the full knowledge of the online parameters is not available a priori, and provide competitive ratio under specific scenarios.

\textbf{Offline fractional model}: In case of fractional revenue model (for offline case), the scheduling problem is more straightforward to tackle since the underlying problem is linear. In this paper, we propose \fcs as an optimal solution for offline fractional case with low time complexity.

\textbf{Online fractional model}: In the online scenario, the scheduling problem is fundamentally more challenging. In our recent study \cite{alinia2018competitive}, 
we proposed two online algorithms referred to as WFAIR and WRAND for EV scheduling problem. The proposed algorithms provide a competitive ratio of $2-1/U$ where $U$ is "scarcity level". However, the result only holds when all EVs have the same maximum charging rate. Also, \cite{alinia2018competitive} does not study the integral revenue model. In this paper, we propose \focs as a 2-competitive online algorithm with no assumption on the input parameters. Moreover, the current study investigates both fractional and integral revenue model while considering total peak constraint in the formulation and addressing multiple charging station scenario. 
In another work~\cite{firstfit}, two simple and natural online algorithms called \textsc{FirstFit} and \textsc{EndFit} were developed and are proved to be 2-competitive. \focs is extension of these algorithms by taking into account the maximum charging rate of the EVs and multiple charging station scenario (see ``Remarks'' at Section IV). As recent studies, \cite{zheng2016online,deng2016whether} provide online algorithms for EV scheduling problem, however, they do not take into account global peak constraint in the underlying problem.

\textbf{Online integral model}: Another direction is to tackle online scenario under integral revenue model. In this setting, the scheduling problem becomes fundamentally more challenging. We note that the integral scheduling problem is strongly NP-Hard even in the offline case \cite{de2018complexity}. Due to combined online and combinatorial challenges, there are very limited studies in this category. \cite{lucier2013efficient} provides an online algorithm without considering the maximum processing rate of the jobs where its competitive ratio can be arbitrary bad depending on ``slackness'' parameter. 
Our work in Section~\ref{sec:onlineint} extends this study for multiple station scenario and considering maximum charging rate for the EVs.
}

@@

\revv{
\subsection{Optimal Offline Design}
\label{sec:fopt}	
\bt{
In this section, we propose an optimal algorithm for offline fractional model. As a natural solution, one may think of a greedy algorithm with the strategy as follows: at each time slot, charge the EV with highest unit-value and process others if only the resource cannot be used by this EV (probably due to maximum charging rate constraint). 
This approach is a popular, yet straightforward method in the online scheduling. However, it turns out that this approach cannot provide an optimal solution to the problem even in fractional revenue model. In fact, this approach only provides a 2-competitive solution. i.e., the profit obtained by the algorithm would be $1/2$ of the optimum. As an intuition of the proof think of two EVs where the one with lower unit value cannot wait due to its deadline and the EV with higher unit value has a long park time. Then, the optimal solution here could be first serving the EV with earlier deadline (and not the one with highest unit values).
}

We refer the proposed algorithm in this secion as the \fcs and summarize it as Algorithm~\ref{alg:optimal-off}.
The \fcs works in two phases. In the first phase (Section~\ref{sec:ph1}), the algorithm decides on the amount of resource to be allocated to each EV within its availability window and reserves resources accordingly. In this phase, the details of allocation is not known. The actual resource allocation is done in the second phase (Section~\ref{sec:ph2}) by setting variables $y_i^t$.

@@

		\begin{algorithm}%\small%1
			\footnotesize
			\caption{\ics}
			\label{alg:ics}
			\DontPrintSemicolon 
%			\KwIn{$n$ EVs with $a_i, d_i, v_i, D_i,$ and $k_i$ associated with each EV $i$, $m$ CSs, local and global peak constraints $p_j, j=1,\dots ,m$, and $p^\mathsf{total}$}

@@

\begin{algorithm}%\small%1
\footnotesize
\caption{\iocs: $\forall t\in \{1,2,\dots ,T\}$}
\label{alg:iocs}
\DontPrintSemicolon 
%\KwIn{$n$ EVs to arrive on the fly, number of CSs $m$, global peak constraint $p^{\textrm{total}}$, local peak constraints $p_j, j=1,\dots ,m$}

### 2019-08-27 TSC/R1/res.tex
BA - Update 2

\ifodd 1
\newcommand{\rev}[1]{{\color{black}#1}}%revise of the text
\newcommand{\com}[1]{\textbf{\color{red}(Bahram says: #1)}}%comment of the text
\newcommand{\comm}[1]{\textbf{\color{red}(Mohammad says: #1)}}%comment of the text
\else
\newcommand{\rev}[1]{#1}
\newcommand{\com}[1]{}
\fi

@@

\begin{enumerate}
\item \textit{...}.

@@

In addition, in the paper, we clearly distinguished between the results in this paper and our previous conference paper on IEEE/ACM IWQoS 2018. Our work in this paper considers more general problem formulation, with global peak constraints, and a comprehensive revenue model. The developed algorithms in Section V, for the integral revenue model, are new in this work. Finally, the trace-driven experimental results are more extensive in this paper and provide comparisons with an algorithm that are currently in use in a real-world charging station.

We highlighted both of the revisions in the revised manuscript and also addressed several other comments that are proposed by the reviewers.

@@

\textit{To clarify this, we added an explanation in Section ...   @Bahram: please add a subsection in Related work and explain IWQoS and other theoretical papers. This is the most important task for this revision. }

@@

$\vartriangleright$ \noindent\textbf{Response:} 
We added an example in Page xx to demonstrate why the simple approach mentioned by the reviewer is not optimal.

{\color{red}@ Mohammad: I copied the following from TSG response letter. We can update and add it to the paper.}

@@

We note that the derived complexities of the algorithms in the paper include parameter $T$ as number of time slots in a day (please see Table II in the manuscript). In large scale scenarios, the effect of increasing $T$ would be negligible in performance of the algorithms as long as it $T/N$ is a constant ($N$ is number of EVs). Therefore, when $N$ grows large and $T$ is a constant, $T$ can be removed from the complexities in the Table II.

@@

$\vartriangleright$ \noindent\textbf{Response:} 
The complexity of the algorithms are low and we just set the charging rate by a simple sorting on arrival of each EV. \textbf{Bahram, please confirm.}

>>>>

\ifodd 1
\newcommand{\bt}[1]{{\color{red}#1}}%revise of the text
\newcommand{\rev}[1]{{\color{black}#1}}%revise of the text
\newcommand{\com}[1]{\textbf{\color{red}(Bahram says: #1)}}%comment of the text
\newcommand{\comm}[1]{\textbf{\color{red}(Mohammad says: #1)}}%comment of the text
\else
\newcommand{\bt}[1]{#1}
\newcommand{\rev}[1]{#1}
\newcommand{\com}[1]{}
\fi

@@

\begin{enumerate}
\item \textit{\color{red} @Mohammad: better to remove this bullet section as we don't have many items to talk about (the only main change is in Related Work)}.

@@

In addition, in the paper, we clearly distinguished between the results in this paper and our previous conference paper on IEEE/ACM IWQoS 2018. Our work in this paper considers more general problem formulation, with global peak constraints, multiple station scenario and a comprehensive revenue model. \bt{Although  ``fractional revenue'' model is studied in both papers (and in some other studies), the proposed algorithms are  different in idea, results and underlying system model. Also, the Section V for the integral revenue model is new in this work. Finally, the trace-driven experimental results are more extensive in this paper and provide comparisons with an algorithm that are currently in use in a real-world charging station. }

%We highlighted both the revision in the revised manuscript and also addressed several other comments that are proposed by the reviewers.

@@

\bt{To clarify this, we added an explanation in last paragraph of the Introduction section and Section II-3 when describing related works.   }

@@

$\vartriangleright$ \noindent\textbf{Response:} 
Thanks for the comment. This approach is equal to running an optimal algorithm at each time slot considering only available EVs and assuming no more EVs will arrive. We added a paragraph in the beginning of Section IV.A to explain this.

{\color{red}@ Mohammad: Should we use the following example in here or in the paper?}

We added an example in Page xx to demonstrate why the simple approach mentioned by the reviewer is not optimal.

@@

We note that the derived complexities of the algorithms in the paper include parameter $T$ as number of time slots in a day (please see Table II in the manuscript). In large scale scenarios, the effect of increasing $T$ would be negligible in performance of the algorithms as long as it $T/N$ is a constant ($n$ is number of EVs). Therefore, when $n$ grows large and $T$ is a constant, $T$ can be removed from the complexities in the Table II.

@@

$\vartriangleright$ \noindent\textbf{Response:} 
\bt{As explained in the previous comment, we derived the time complexity of all proposed algorithms (listed in Table II in the paper) which indicate their running time in large scale scenarios (i.e., when $n$ is large). If we assume that $T$ is a constant (which is a quite realistic assumption), the complexity of all algorithms simplifies to $O(n^2)$. Hence, we believe that a normal PC could be fast enough to run the algorithms.}
